---
title: "Homework 2"
output: html_notebook
---

# Unsupervised Learning

### Find model parameters using EM 
The dataset below contains eruption data from old faithful.  
Analysts note that old faithful erupts in certain patterns: Sometimes there are long eruptions, sometimes there are short.
The eruptions are also followed by a delay that can vary accordingly.

Read in the data to get started.


```{r}
data(faithful)
# investigate the old faithful eruption patterns using scatter plots, etc.  
```

## Q1: How many clusters do you believe exist?

## Q2: How did you arrive at your conclusion?

## Q3 : Use Expectation Maximization clustering 
Extract cluster parameters using your chosen``k`` cluster count  and report them below
```{r}
library(EMCluster)
# shortemcluster(faithful, simple.init(faithful, nclass = k))
```


## Q4: Use dbscan to perform clustering.  

```{r}
library(dbscan)
# dbscan(faithful, minPts, eps)

```
Report the settings you chose for minPts and epsilon, and how you arrived at them (hint: histograms and distances)

## Q5: Use kmeans to perform clustering
Use the k you chose before

```{r}
# kmeans(faithful, k) should already be available in your R environment
```

## Q6 : Which clustering technique works the best here?  
